More equivalences for Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2018-07-03 2016-Weak-FH-Bisim/WeakBisim.tex
orga section weak bisim + theorems + decomp lemmas in appendix

Finally, the previous theorems can be composed to state a general theorem about 
composability and FH-bisimilarity.
\begin{corollary}[Composability]
	Consider two FH-bisimilar pNets with an arbitrary number of holes, when replacing, 
	inside those two original pNets, a subset of the holes by FH-bisimilar pNets, we 
	obtain two FH-bisimilar pNets.
\end{corollary}
%
%In the following we use $\approx$ to denote equality modulo alpha conversion of free 
%variables. In practice, $x\approx y$ can be obtained if: $x=clone(a) \land y=clone(a)$.

@@

@@

--------------------------------------------------------

@@

@@

\TODO{We will discuss this corollary and the title.}
\begin{corollary}[Weak Composability]
	Consider two weak FH-bisimilar pNets with an arbitrary number of holes, when replacing, 
	inside those two original pNets, a subset of the holes by weak FH-bisimilar pNets, we 
	obtain two weak FH-bisimilar pNets.
\end{corollary}

@@

       \section{Weak FH-bisimulation lemmas and proofs}
\begin{lemma}[Weak OT decomposition]\label{lem-decompose} 
	Let $\Leaves(Q)=p_l^{l\in L_Q}$; suppose\footnote{Note that the hypotheses of the 
	lemma imply that $Q$ is 
	not a \pLTS\ but a similar lemma can be proven for a pLTS $Q$}:
	\[ P[Q]_{j_0}  
		\models
		{\openrule
			{
				\gamma_j^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTWeakarrow {\alpha}
				\ostate{s_i'^{\, i\in L}}}
		}
	\]
		with  $J\cap\Holes(Q)\neq\emptyset$ or $\exists i\in L_Q.\,s_i\neq s'_i$, i.e. $Q$ takes part in the reduction.
		 Then there exist $v_Q$, $\Pred\,'$,  
		$\Post\,'$,   $\gamma_{j_0}=[\beta_0..\beta_n]$, and for all $p\in[0..n]$ there exist $\Pred_p$, $\Post_p$ and a family $\gamma_{p j}^{j\in J_p}$ and $s_{p i}^{\,i\in L_Q}$. s.t.:\\[-2ex]
		%\[
		\begin{mathpar}
		P\models{\openrule
			{
				\gamma_j^{j\in (J_p\setminus \Holes(Q)) \cup \{j_0\}}, 
				\Pred\,',  
				\Post\,'}
			{\ostate{s_i^{i\in L\setminus L_Q}} \OTWeakarrow {\alpha}
				\ostate{s_i'^{\,i\in L\setminus L_Q}}}
		}%\]
	\vspace{-2.2ex}\\\text{and for all $p\in[0..n]$~~}
		%\[
		Q\models{\openrule
			{
				\gamma_{p j}^{j\in J_p}, \Pred_p,  
				\Post_p}
			{\ostate{s_i^{i\in L_Q}} \OTWeakarrow {\alpha_i}
				\ostate{s_{p i}^{\,i\in L_Q}}}
		}%\]
\\
		\text{and~~}  \bigcup_{\begin{array}{l}
p\in[0..n]\\{j\in J_p}
\end{array}} J_{p j} = J\cap\Holes(Q) \qquad
 \text{and\quad} \forall i \in L_Q.\, s_{n i} = s_i'\\{\hspace{-3ex}}\text{and~~} \Pred \iff \Pred\,'
		\land \!\!\bigwedge_{p\in[0..n]}(\alpha_p=\beta_p\land \Pred_p) \qquad\text{and~}
\Post=\Post\,'\uplus\!\! \biguplus_{p\in[0..n]}
		\Post_p  
		\end{mathpar}
where $\Post_p$ only act upon variables of 
		$\Leaves(Q)$.
\end{lemma}

>>>>

Finally, the previous theorems can be composed to state a general theorem about 
composability and FH-bisimilarity.
\begin{theorem}[Composability]
	Consider two FH-bisimilar pNets with an arbitrary number of holes, when replacing, 
	inside those two original pNets, a subset of the holes by FH-bisimilar pNets, we 
	obtain two FH-bisimilar pNets.
\end{theorem}
%
%In the following we use $\approx$ to denote equality modulo alpha conversion of free 
%variables. In practice, $x\approx y$ can be obtained if: $x=clone(a) \land y=clone(a)$.

@@

\subsection{Defining weak open transitions}

@@

\begin{theorem}[Weak FH-Bisimulation is an equivalence]\label{thm-weak-equiv} Suppose $\mathcal{R}$ 
is a weak FH-bisimualtion. Then $\mathcal{R}$ is an equivalence, that is, $\mathcal{R}$ is 
reflexive, symmetric and transitive.
\end{theorem}
proof in appendix

@@

\TODO{question : we could merge prop 1 and def above, is it better?}

@@

\begin{theorem}[Congruence for  weak bisimulation]\label{weak-thm-congr-eq}
	Consider an open pNet:
	$\pNet = \mylangle \pNet_i^{i\in I}, \Sort_j^{j\in J}, 
	\overline{\symb{SV}}\myrangle$.
	Let $j_0\in J$ be a hole. Let $Q$ and $Q'$ be two weak FH-bisimilar pNets such that 
	$\Sortop(Q)=\Sortop(Q')=\Sort_{j_0}$\footnote{Note that $\Sortop(Q)=\Sortop(Q')$ is 
	ensured by 
	strong bisimilarity.}. Then 
	$\pNet[Q]_{j_0}$ and 
	$\pNet[Q']_{j_0}$ are weak FH-bisimilar.
\end{theorem}

\begin{theorem}[Context equivalencefor  weak bisimulation]\label{weak-thm-ctxt-eq}
	Consider two weak FH-bisimilar open pNets:
	$\pNet = \mylangle \pNet_i^{i\in I}, \Sort_j^{j\in J}, 
	\overline{\symb{SV}}\myrangle$ and 	$\pNet' = \mylangle {\pNet'}_i^{i\in I}, 
	\Sort_j^{j\in 
	J}, 	\overline{\symb{SV'}}\myrangle$ 
	(recall they must have the same holes to be bisimilar).
	Let $j_0\in J$ be a hole, and $Q$ be a pNet such that $\Sortop(Q)=\Sort_{j_0}$. Then 
	$\pNet[Q]_{j_0}$ and 
	$\pNet'[Q]_{j_0}$ are weak FH-bisimilar.
\end{theorem}

Finally, the previous theorems can be composed to state a general theorem about 
composability and weak FH-bisimilarity.

\begin{theorem}[Composability of weak bisimulation]\label{weak-compos}
	Consider two weak FH-bisimilar pNets with an arbitrary number of holes, when replacing, 
	inside those two original pNets, a subset of the holes by weak FH-bisimilar pNets, we 
	obtain two weak FH-bisimilar pNets.
\end{theorem}

@@

       \section{Weak FH-bisimulation lemmas and proofs}

\subsection{Weak bisimulation is an equivalence}
\TODO{moved from body}

\subsection{composition properties}
\TODO{Ludo: in the next 2 theorems I have a doubt about the biguplus of post p : not sure they should be disjoint, but else we need to do a complex composition  of posts}
\begin{lemma}[Weak OT decomposition]\label{lem-decompose} 
	Let $\Leaves(Q)=p_l^{l\in L_Q}$; suppose\footnote{Note that the hypotheses of the 
	lemma imply that $Q$ is 
	not a \pLTS\ but a similar lemma can be proven for a pLTS $Q$}:
	\[ P[Q]_{j_0}  
		\models
		{\openrule
			{
				\gamma_j^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTWeakarrow {\alpha}
				\ostate{s_i'^{\, i\in L}}}
		}
	\]
		with  $J\cap\Holes(Q)\neq\emptyset$ or $\exists i\in L_Q.\,s_i\neq s'_i$, i.e. $Q$ takes part in the reduction.
		 Then there exist $v_Q$, $\Pred\,'$,  
		$\Post\,'$,   $\gamma_{j_0}=[\beta_0..\beta_n]$, and for all $p\in[0..n]$ there exist $\Pred_p$, $\Post_p$ and a family $\gamma_{p j}^{j\in J_p}$ and for all $p\in[0..n+1]$ $s_{p i}^{\,i\in L_Q}$. s.t.:\\[-2ex]
		%\[
		\begin{mathpar}
		P\models{\openrule
			{
				\gamma_j^{j\in (J_p\setminus \Holes(Q)) \cup \{j_0\}}, 
				\Pred\,',  
				\Post\,'}
			{\ostate{s_i^{i\in L\setminus L_Q}} \OTWeakarrow {\alpha}
				\ostate{s_i'^{\,i\in L\setminus L_Q}}}
		}%\]
	\vspace{-2.2ex}\\\text{and for all $p\in[0..n]$~~}
		%\[
		Q\models{\openrule
			{
				\gamma_{p j}^{j\in J_p}, \Pred_p,  
				\Post_p}
			{\ostate{s_{p i}^{i\in L_Q}} \OTWeakarrow {\alpha_p}
				\ostate{s_{(p+1) i}^{\,i\in L_Q}}}
		}%\]
\\
		\text{and~~}  \bigcup_{\begin{array}{l}
p\in[0..n]
\end{array}} J_{p} = J\cap\Holes(Q) \qquad
 \text{and\qquad} \forall j\in J_p,\,\gamma_j= \mathop{\dot{\bigcup}}_{p\in [0..n]} \gamma_{p j} \\
 \text{and\quad} \forall i \in L_Q.\, s_{(n+1) i} = s_i'\land s_{0 i} = s_i\\{\hspace{-3ex}}\text{and~~} \Pred \iff \Pred\,'
		\land \!\!\bigwedge_{p\in[0..n]}(\alpha_p=\beta_p\land \Pred_p) \qquad\text{and~}
\Post=\Post\,'\uplus\!\! \biguplus_{p\in[0..n]}
		\Post_p  
		\end{mathpar}
where $\Post_p$ only act upon variables of 
		$\Leaves(Q)$.
\end{lemma}
\TODO{proof}

\begin{lemma}[Weak open transition composition]\label{lem-compose} 
	Suppose $j_0\in J$ and $\gamma_{j_0}=[\beta_0..\beta_n]$ and additionally:\\[-2ex]
\begin{mathpar}
%\[
P\models{\openrule
	{
		\gamma_j^{j\in J}, 
		\Pred,  
		\Post}
	{\ostate{s_i^{i\in L}} \OTWeakarrow {\alpha}
		\ostate{s_i'^{\, i\in L}}}
}\\
\text{~~and for all $p\in[0..n]$~~}
Q\models{\openrule
	{
		\gamma_{p j}^{j\in J_p},
		 \Pred_p,  
		\Post_p}
	{\ostate{s_{p i}^{i\in L_p}} \OTWeakarrow {\alpha_p}
		\ostate{s_{(p+1) i}^{\, i\in L_Q}}}
}%\]
\end{mathpar}
Let 
\begin{mathpar}
J_Q=\bigcup_{\begin{array}{l}
p\in[0..n]
\end{array}} J_{p}
\text{\qquad and\qquad} 
\Pred\,'=\Pred\land   \!\!\bigwedge_{p\in[0..n]}(\alpha_p=\beta_p\land \Pred_p)
\\
\forall j\in J_p,\,\gamma_j= \mathop{\dot{\bigcup}}_{p\in [0..n]} \gamma_{p j}
\end{mathpar}
Then, we have\\[-2ex]
	\[ P[Q]_{j_0}  
	\models
	{\openrule
		{
			\gamma_j^{(j\in J\setminus\{j_0\}) \uplus J_Q}, 
			\Pred\,',  
			\Post\uplus\biguplus_{p\in[0..n]}
		\Post_p  }
		{\ostate{s_i^{i\in L\uplus L_Q}} \OTWeakarrow {\alpha}
			\ostate{s_i'^{\, i\in L\uplus L_Q}}}
	}
	\]
\end{lemma}
\TODO{proof}

\TODO{prove the 3 theorems for composability of weak bisim or explain why the proof is similar to strong (thanks to the previous lemmas)}
